Fast Nearest Neighbour Classification

### 2016-05-10 paper/paper.tex
semifinish

@@

@@

@@

As described above, the following benchmarks will not compare the runtime of the mentioned algorithms. The benchmarks
will compare the calls of the distance function that the algorithm needs for preprocessing and for finding a nearest
neighbour. Those benchmarks can make sense when the distance function has a heavy complexity. For example the string
metric Levenshtein distance or in higher dimensional vector spaces.\\
In general a benchmark is affected by the implementation, by the executing machine, by the domain of the problem
and the distribution of the example. As the following benchmarks only counting the calls of the distance function, they
aren't affected by the implementation or the executing machine.

@@

The algorithms above are all implemented in Java and can be found in the following git repository\\
\url{https://github.com/GordonLesti/FastNearestNeighbourClassification}.\\

A nearest neighbour algorithm implemented in Java should handle every type of object that a set $\mathbb{U}$ can
represent under the condition that there is a metric on $\mathbb{U}$. Because of that, a nearest neighbour algorithm
class needs to extend from the abstract class\\ \verb+FastNearestNeighbourClassificator<T, D extends Comparable<D>>+
that can be found in the package \verb+fnnc.algo+. The generic \verb+T+ represents an object from the set $\mathbb{U}$
and the generic \verb+D+ represents the distance and has to be of type \verb+Comparable+ from the package
\verb+java.lang+.\\

@@

The abstract class\\ \verb+FastNearestNeighbourClassificator<T, D extends Comparable<D>>+ has two protected class
variables:
\begin{itemize}
	\item a \verb+distanceCalculator+ of type \verb+DistanceCalculator<T, D>+
	\item a \verb+objectCollection+ of type \verb+Collection<T>+
\end{itemize}
and three simple class methods:
\begin{itemize}
	\item a constructor that takes an object of type \verb+DistanceCalculator<T, D>+ from the package
		\verb+fnnc.model+ and sets the protected class variable
	\item a default function \verb+preProcessing+ that takes a \verb+Collection<T>+ from the package
		\verb+java.util+ and sets the protected class variable
	\item a abstract function \verb+calculateNearestNeighbour+ that takes the query object of type \verb+T+ and
		returns the nearest neighbour object of type \verb+T+
\end{itemize}

A instance of type \verb+FastNearestNeighbourClassificator+ shouldn't and can't calculate the distance between two
objects. So that a instance of \verb+DistanceCalculator+ should be injected into the constructor. All implemented
algorithms exclude the \verb+FullSearch+ require that the \verb+DistanceCalculator+ returns a distance of type
\verb+Double+ instead of only \verb+Comparable+. This is connected with the fact that only the Full Search can operate
in a not metric space.

@@

The abstract class\\ \verb+DistanceCalculator<Point2D.Double, Double>+ has just one private \verb+count+ of type
\verb+int+ and three class methods:
\begin{itemize}
	\item a public final function \verb+calculateDistance+ that takes two objects of type \verb+T+, increases the
		internal count and returns the result of \verb+internalCalculateDistance+
	\item a public final function \verb+reset+ that resets the internal count
	\item a protected abstract function \verb+internalCalculateDistance+ that takes two objects of type \verb+T+ and
		returns the distance object of type \verb+D+
\end{itemize}

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{ | l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	$n$ & 0 & $\frac{n(n-1)}{2}$ & $\frac{n(n-1)}{2}$ & $n$ & $\frac{n(n-1)}{2}$ \\ \hline
		\end{tabular}
	\end{center}
	\caption{The calls of the distance function $D$ in the preprocessing depending on the size $n$ of the test data set
		$S$.}
	\label{fig:preprocessing:tabular}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{ | l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 100 & 29 & 20 & 46 & 4 \\ \hline
			200 & 200 & 39 &26 & 82 & 4 \\ \hline
			300 & 300 & 46 & 30 & 117 & 4 \\ \hline
			400 & 400 & 53 & 35 & 160 & 4 \\ \hline
			500 & 500 & 57 & 40 & 189 & 4 \\ \hline
			600 & 600 & 63 & 42 & 228 & 4 \\ \hline
			700 & 700 & 68 & 44 & 263 & 4 \\ \hline
			800 & 800 & 73 & 46 & 303 & 4 \\ \hline
			900 & 900 & 78 & 50 & 341 & 4 \\ \hline
			1000 & 1000 & 79 & 51 & 367 & 4 \\ \hline
			1100 & 1100 & 84 & 56 & 409 & 4 \\ \hline
			1200 & 1200 & 86 & 56 & 442 & 4 \\ \hline
			1300 & 1300 & 89 & 59 & 466 & 4 \\ \hline
			1400 & 1400 & 90 & 60 & 510 & 4 \\ \hline
			1500 & 1500 & 97 & 63 & 541 & 4 \\ \hline
			1600 & 1600 & 100 & 66 & 573 & 4 \\ \hline
			1700 & 1700 & 102 & 67 & 624 & 4 \\ \hline
			1800 & 1800 & 104 & 69 & 660 & 4 \\ \hline
			1900 & 1900 & 109 & 71 & 664 & 4 \\ \hline
			2000 & 2000 & 113 & 73 & 740 & 4 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The average calls of the distance function $D$ in the query processing depending on the size of the test
		data set $S$.}
	\label{fig:queryprocessing:tabular}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{ | l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 0 & 13 & 8 & 23 & 1 \\ \hline
			200 & 0 & 17 & 11 & 46 & 1 \\ \hline
			300 & 0 & 21 & 13 & 69 & 1 \\ \hline
			400 & 0 & 25 & 15 & 94 & 1 \\ \hline
			500 & 0 & 27 & 17 & 119 & 1 \\ \hline
			600 & 0 & 29 & 19 & 144 & 1 \\ \hline
			700 & 0 & 32 & 19 & 162 & 1 \\ \hline
			800 & 0 & 33 & 21 & 190 & 1 \\ \hline
			900 & 0 & 36 & 22 & 213 & 1 \\ \hline
			1000 & 0 & 37 & 23 & 232 & 1 \\ \hline
			1100 & 0 & 38 & 24 & 264 & 1 \\ \hline
			1200 & 0 & 40 & 26 & 277 & 1 \\ \hline
			1300 & 0 & 43 & 27 & 301 & 1 \\ \hline
			1400 & 0 & 43 & 27 & 339 & 1 \\ \hline
			1500 & 0 & 47 & 27 & 350 & 1 \\ \hline
			1600 & 0 & 46 & 29 & 374 & 1 \\ \hline
			1700 & 0 & 48 & 29 & 401 & 1 \\ \hline
			1800 & 0 & 49 & 31 & 429 & 1 \\ \hline
			1900 & 0 & 52 & 32 & 442 & 1 \\ \hline
			2000 & 0 & 52 & 33 & 487 & 1 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The standard deviation of the calls of the distance function $D$ in the query processing depending on the
		size of the test data set $S$.}
	\label{fig:standarddeviation:tabular}
\end{figure}

@@

>>>>

\paragraph{Preprocessing}

@@

\paragraph{Query Processing}

@@

\subsubsection{Orchard’s Algorithm with marked bits}

There exists a improved version of the Orchard’s Algorithm to ensure that no distance between an item $x \in S$ and $q$
is calculated twice. This can achieved by boolean flags for every item in the query processing of the Orchard’s
Algorithm.

@@

As described above, the following benchmarks will not compare the runtime of the mentioned algorithms. The benchmarks
will compare the calls of the distance function that the algorithm needs for preprocessing and for finding a nearest
neighbour. Those benchmarks can make sense when the distance function has a heavy complexity. For example the string
metric Levenshtein distance or in higher dimensional vector spaces.\\
In general a benchmark is affected by the implementation, by the programming language, by the executing machine, by the
domain of the problem and the distribution of the example. As the following benchmarks only counting the calls of the
distance function, they are not affected by the programming language or the executing machine.

@@

The algorithms above are all implemented in Java and can be found in the following git
repository\footnote[1]{\url{https://github.com/GordonLesti/FastNearestNeighbourClassification}}.\\
Since the benchmakrs are independed from the programming language and the executing machine, they use sometimes not the
most effective data structures, but that never effects the logical flow of the algorithms.\\

A nearest neighbour algorithm implemented in Java should handle every type of object that a set $\mathbb{U}$ can
represent under the condition that there is a metric on $\mathbb{U}$. Because of that, a nearest neighbour algorithm
class needs to extend from the abstract class\\ \verb+FastNearestNeighbourClassificator<T, D extends Comparable<D>>+
that\\ can be found in the package \verb+fnnc.algo+. The generic \verb+T+ represents an object from the set $\mathbb{U}$
and the generic \verb+D+ represents the distance and has to be of type \verb+Comparable+ from the package
\verb+java.lang+.\\

@@

The abstract class\\ \verb+FastNearestNeighbourClassificator<T, D extends Comparable<D>>+ has\\ two protected class
variables:
\begin{itemize}
	\item a \verb+distanceCalculator+ of type \verb+DistanceCalculator<T, D>+
	\item a \verb+objectCollection+ of type \verb+Collection<T>+
\end{itemize}
and three simple class methods:
\begin{itemize}
	\item a constructor that takes an object of type \verb+DistanceCalculator<T, D>+ from the package
		\verb+fnnc.model+ and sets the protected class variable
	\item a default function \verb+preProcessing+ that takes a \verb+Collection<T>+ from the package
		\verb+java.util+ and sets the protected class variable
	\item a abstract function \verb+calculateNearestNeighbour+ that takes the query object of type \verb+T+ and
		returns the nearest neighbour object of type \verb+T+
\end{itemize}

A instance of type \verb+FastNearestNeighbourClassificator+ should not be allowed and can not calculate the distance
between two objects. So that a instance of \verb+DistanceCalculator+ should be injected into the constructor. All
implemented algorithms exclude the \verb+FullSearch+ require that the\\ \verb+DistanceCalculator+ returns a distance of
type \verb+Double+ instead of only\\ \verb+Comparable+. This is connected with the fact that only the Full Search can
operate in a not metric space.

@@

The abstract class\\ \verb+DistanceCalculator<Point2D.Double, Double>+ has just one private \verb+count+ of type
\verb+int+ and three class methods:
\begin{itemize}
	\item a public final function \verb+calculateDistance+ that takes two objects of type \verb+T+, increases the
		internal count and returns the result of\\ \verb+internalCalculateDistance+
	\item a public final function \verb+reset+ that resets the internal count
	\item a protected abstract function \verb+internalCalculateDistance+ that takes two objects of type \verb+T+ and
		returns the distance object of type \verb+D+
\end{itemize}

@@

The results of the preprocessing are static. Figure~\ref{fig:preprocessing:tabular} and
figure~\ref{fig:preprocessing:graph} summarizing again the complexity of the preprocessing for every algorithm
depending on the size of set $S$.

\begin{figure}
	\begin{center}
		\begin{tabular}{| l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	$n$ & 0 & $\frac{n(n-1)}{2}$ & $\frac{n(n-1)}{2}$ & $n$ & $\frac{n(n-1)}{2}$ \\ \hline
		\end{tabular}
	\end{center}
	\caption{The calls of the distance function $D$ in the preprocessing depending on the size $n$ of the test data set
		$S$.}
	\label{fig:preprocessing:tabular}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{| l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 100 & 29 & 20 & 46 & 4 \\ \hline
			200 & 200 & 39 &26 & 82 & 4 \\ \hline
			300 & 300 & 46 & 30 & 117 & 4 \\ \hline
			400 & 400 & 53 & 35 & 160 & 4 \\ \hline
			500 & 500 & 57 & 40 & 189 & 4 \\ \hline
			600 & 600 & 63 & 42 & 228 & 4 \\ \hline
			700 & 700 & 68 & 44 & 263 & 4 \\ \hline
			800 & 800 & 73 & 46 & 303 & 4 \\ \hline
			900 & 900 & 78 & 50 & 341 & 4 \\ \hline
			1000 & 1000 & 79 & 51 & 367 & 4 \\ \hline
			1100 & 1100 & 84 & 56 & 409 & 4 \\ \hline
			1200 & 1200 & 86 & 56 & 442 & 4 \\ \hline
			1300 & 1300 & 89 & 59 & 466 & 4 \\ \hline
			1400 & 1400 & 90 & 60 & 510 & 4 \\ \hline
			1500 & 1500 & 97 & 63 & 541 & 4 \\ \hline
			1600 & 1600 & 100 & 66 & 573 & 4 \\ \hline
			1700 & 1700 & 102 & 67 & 624 & 4 \\ \hline
			1800 & 1800 & 104 & 69 & 660 & 4 \\ \hline
			1900 & 1900 & 109 & 71 & 664 & 4 \\ \hline
			2000 & 2000 & 113 & 73 & 740 & 4 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The average calls of the distance function $D$ in the query processing depending on the size of the test
		data set $S$.}
	\label{fig:queryprocessing:tabular}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{| l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 0 & 13 & 8 & 23 & 1 \\ \hline
			200 & 0 & 17 & 11 & 46 & 1 \\ \hline
			300 & 0 & 21 & 13 & 69 & 1 \\ \hline
			400 & 0 & 25 & 15 & 94 & 1 \\ \hline
			500 & 0 & 27 & 17 & 119 & 1 \\ \hline
			600 & 0 & 29 & 19 & 144 & 1 \\ \hline
			700 & 0 & 32 & 19 & 162 & 1 \\ \hline
			800 & 0 & 33 & 21 & 190 & 1 \\ \hline
			900 & 0 & 36 & 22 & 213 & 1 \\ \hline
			1000 & 0 & 37 & 23 & 232 & 1 \\ \hline
			1100 & 0 & 38 & 24 & 264 & 1 \\ \hline
			1200 & 0 & 40 & 26 & 277 & 1 \\ \hline
			1300 & 0 & 43 & 27 & 301 & 1 \\ \hline
			1400 & 0 & 43 & 27 & 339 & 1 \\ \hline
			1500 & 0 & 47 & 27 & 350 & 1 \\ \hline
			1600 & 0 & 46 & 29 & 374 & 1 \\ \hline
			1700 & 0 & 48 & 29 & 401 & 1 \\ \hline
			1800 & 0 & 49 & 31 & 429 & 1 \\ \hline
			1900 & 0 & 52 & 32 & 442 & 1 \\ \hline
			2000 & 0 & 52 & 33 & 487 & 1 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The standard deviation of the calls of the distance function $D$ in the query processing depending on the
		size of the test data set $S$.}
	\label{fig:standarddeviation:tabular}
\end{figure}

@@

\section{Conclusions}

\paragraph{Full Search} The results may leave the impression that the Full Search is not a good choice. But this
algorithm remains useful if the set $S$ changes completely or when $\mathbb{U}$ is no metric space.

\paragraph{Annulus Method} The Annulus Method has a very complex query processing that

\paragraph{Orchard’s Algorithm with and without marking bits} In the query processing the Orchard’s Algorithm with
marking bits consumed on average only 66\% of the distance function calls that the algorithm needed without marking
bits.

\section{Future work}

The LAESA algorithm and Metric Trees are missing in the benchmarks above. Both approaches are promising and their
results compared to algorithms of this paper would be interesting.
